### 2023-03-25 index.qmd
Quarto-ify the project

---
title: "The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society"
author:
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@student.gsu.edu
  orcid: 0000-0002-6552-8338
format:
  html:
    code-tools: false
    toc: false
---

## Abstract

Abstract goes here

## Code

All the raw code and data for this paper is available in a [GitHub repository](https://github.com/andrewheiss/compassionate-clam).