### 2023-03-25 notebook/manuscript.qmd

---
title: Manuscript
toc: false
page-layout: full
---

The final manuscript will be downloadable from here. ([See this for a similar example.](https://stats.andrewheiss.com/mountainous-mackerel/analysis/manuscript.html))