### 2023-03-28 notebook/exploratory.qmd
Process map data with a target + make map

---
title: EDA
---

# Map!

```{r load-packages-data, message=FALSE, warning=FALSE}
library(tidyverse)
library(targets)
library(here)
library(sf)

# Point to the _targets folder location since this qmd is in a subfolder
tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

# Load targets
mapdata <- tar_read(ongo_mapdata)
```

```{r ongo-map, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
labeldata <- mapdata |> 
  filter(ro_count > 0)

plot_map <- ggplot() +
  geom_sf(data = mapdata,
          aes(fill = ro_count), color = "white") +
  scale_fill_viridis_c(option = "mako", begin = 0.1, end = 0.9, direction = -1) +
  geom_sf_label(data = labeldata, aes(label = ro_count), 
                nudge_y = -0.8 , nudge_x = -0.1, size = 3, alpha = 0.7) +
  theme_void()
plot_map
```