### 2023-04-06 notebook/analysis-local.qmd
Create targets for models + preliminary results

---
title: "Local connections"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(brms)
library(marginaleffects)
library(tidybayes)
library(ggdist)
library(patchwork)
library(scales)

# Generated via random.org
set.seed(196491)

# Load targets
tar_load(ongo)
tar_load(m_full_ordbeta)
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))

prop_to_provinces <- function(x, lower = 1, upper = 32) {
  (x * (upper - lower)) + lower
}
```

```{r generate-preds}
preds <- m_full_ordbeta |> 
  predicted_draws(datagrid(model = m_full_ordbeta, local_connect = unique), 
                  ndraws = 1000)

epreds <- m_full_ordbeta |> 
  epred_draws(datagrid(model = m_full_ordbeta, local_connect = unique), 
              ndraws = 1000)
```

```{r plot-epred-local}
epreds |> 
  ungroup() |> 
  mutate(local_connect = factor(local_connect, labels = c("No local connections", "Local connections"))) |> 
  ggplot(aes(x = .epred, y = local_connect)) +
  stat_halfeye(fill = clrs[2]) +
  scale_x_continuous(labels = label_percent(),
                     sec.axis = sec_axis(trans = ~ prop_to_provinces(.),
                                         name = "Expected provinces",
                                         breaks = seq(8, 32, 4))) +
  guides(fill = "none") +
  labs(x = "Expected proportion of provinces", y = NULL) +
  theme_ongo() +
  theme(axis.ticks.x.top = element_line(linewidth = 0.25, color = "grey50"))
```

TODO: Prop of 1s, 32s, and in-betweens