### 2023-04-07 notebook/analysis-local.qmd
Plot discrete and continuous predictions across groups

---
title: "Local connections"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(brms)
library(marginaleffects)
library(tidybayes)
library(ggdist)
library(patchwork)
library(scales)

# Generated via random.org
set.seed(196491)

# Load targets
tar_load(ongo)
tar_load(m_full_ordbeta)
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))

prop_to_provinces <- function(x, lower = 1, upper = 32) {
  (x * (upper - lower)) + lower
}
```

```{r generate-preds}
preds <- m_full_ordbeta |> 
  predicted_draws(datagrid(model = m_full_ordbeta, local_connect = unique), 
                  ndraws = 1000)

epreds <- m_full_ordbeta |> 
  epred_draws(datagrid(model = m_full_ordbeta, local_connect = unique), 
              ndraws = 1000)
```

```{r plot-epred-local}
epreds |> 
  ungroup() |> 
  mutate(local_connect = factor(local_connect, labels = c("No local connections", "Local connections"))) |> 
  ggplot(aes(x = .epred, y = local_connect)) +
  stat_halfeye(fill = clrs[2]) +
  scale_x_continuous(labels = label_percent(),
                     sec.axis = sec_axis(trans = ~ prop_to_provinces(.),
                                         name = "Expected provinces",
                                         breaks = seq(8, 32, 4))) +
  guides(fill = "none") +
  labs(x = "Expected proportion of provinces", y = NULL) +
  theme_ongo() +
  theme(axis.ticks.x.top = element_line(linewidth = 0.25, color = "grey50"))
```

```{r plot-pred-discrete-continuous-local, message=FALSE}
preds_local_plot <- preds %>% 
  ungroup() %>% 
  mutate(local_connect = factor(local_connect, labels = c("No local connections", "Local connections"))) |> 
  mutate(outcome = case_when(
    .prediction == 0 ~ "1 province",
    .prediction > 0 & .prediction < 1 ~ "1–32 provinces",
    .prediction == 1 ~ "32 provinces"
  )) %>% 
  group_by(outcome, local_connect) %>% 
  summarize(count = n()) |> 
  group_by(local_connect) |> 
  mutate(prop = count / sum(count))

ggplot(preds_local_plot, aes(x = local_connect, y = prop)) +
  geom_segment(aes(xend = local_connect, yend = 0), color = "grey50") +
  geom_point(color = clrs[2], size = 2) +
  expand_limits(y = 0) +
  facet_wrap(vars(outcome)) +
  scale_x_discrete(labels = label_wrap(10)) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = NULL, y = "Proportion of predictions") +
  theme_ongo()
```