### 2023-04-09 notebook/analysis-timing.qmd
stop caching now that I'm making the predictions right

---
title: "Registration timing"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(brms)
library(marginaleffects)
library(tidybayes)
library(ggdist)
library(patchwork)
library(scales)

# Generated via random.org
set.seed(196491)

# Load targets
tar_load(ongo)
tar_load(c(m_full_ordbeta, preds_timing))
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))

prop_to_provinces <- function(x, lower = 1, upper = 32) {
  (x * (upper - lower)) + lower
}
```

```{r plot-epred-timing}
epreds <- m_full_ordbeta |> 
  epred_draws(datagrid(model = m_full_ordbeta, years_since_law = 0:5), 
              ndraws = 1000)

epreds |> 
  mutate(years_since_law = factor(years_since_law)) |> 
  ggplot(aes(x = .epred, y = years_since_law)) +
  stat_halfeye(fill = clrs[7]) +
  scale_x_continuous(labels = label_percent(),
                     sec.axis = sec_axis(trans = ~ prop_to_provinces(.),
                                         name = "Expected provinces",
                                         breaks = seq(8, 32, 4))) +
  guides(fill = "none") +
  labs(x = "Expected proportion of provinces", y = "Years since law took effect in 2017") +
  theme_ongo() +
  theme(axis.ticks.x.top = element_line(linewidth = 0.25, color = "grey50"))
```

```{r plot-epred-timing-lineribbon}
epreds %>% 
  ggplot(aes(x = years_since_law, y = .epred)) +
  scale_x_reverse() +
  scale_y_continuous(labels = label_percent(),
                     sec.axis = sec_axis(trans = ~ prop_to_provinces(.),
                                         name = "Expected provinces")) +
  labs(x = "Years since law took effect (2017)", 
       y = "Expected proportion of provinces") +
  stat_lineribbon(alpha = 0.35, fill = clrs[7], color = clrs[7]) +
  theme_ongo() +
  theme(axis.title.y.right = element_text(hjust = 1, margin = margin(l = 10), angle = 90))
```

```{r plot-pred-discrete-continuous-timing, message=FALSE}
preds_timing_plot <- preds_timing |> 
  unnest(preds) |> 
  mutate(outcome = case_when(
    .prediction == 0 ~ "1 province",
    .prediction > 0 & .prediction < 1 ~ "1–32 provinces",
    .prediction == 1 ~ "32 provinces"
  )) %>% 
  group_by(outcome, years_since_law, draw) %>% 
  summarize(count = n()) |> 
  group_by(years_since_law) |> 
  mutate(prop = count / sum(count) * 100)

ggplot(preds_timing_plot, aes(x = years_since_law, y = prop)) +
  stat_lineribbon(alpha = 0.35, fill = clrs[7], color = clrs[7]) +
  expand_limits(y = 0) +
  facet_wrap(vars(outcome)) +
  scale_x_reverse(breaks = c(0:5)) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = "Years since law took effect (2017)", 
       y = "Proportion of predictions") +
  theme_ongo()
```