### 2023-09-13 notebook/model-details.qmd

---
title: Model details
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(brms)
library(marginaleffects)
library(tidybayes)
library(ggdist)
library(patchwork)
library(scales)

# Generated via random.org
set.seed(196491)

# Load targets
tar_load(ongo)
tar_load(m_full_ordbeta)
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))

prop_to_provinces <- function(x, lower = 1, upper = 32) {
  (x * (upper - lower)) + lower
}
```

TODO: Formal definition and justification for ordered beta instead of ZOIB

```{r build-posterior-preds}
full_posterior <- predicted_draws(m_full_ordbeta, 
                                  newdata = m_full_ordbeta$data, 
                                  ndraws = 100)

posterior_categories <- full_posterior |> 
  ungroup() |> 
  mutate(outcome = case_when(
    .prediction == 0 ~ "1",
    .prediction > 0 & .prediction < 1 ~ "1–32",
    .prediction == 1 ~ "32"
  )) |> 
  group_by(outcome, .draw) |> 
  summarize(count = n()) |> 
  ungroup()

actual_counts <- m_full_ordbeta$data |> 
  mutate(outcome = case_when(
    province_count == 0 ~ "1",
    province_count > 0 & province_count < 1 ~ "1–32",
    province_count == 1 ~ "32"
  )) |> 
  group_by(outcome) |> 
  summarize(count = n())

plot_categories <- posterior_categories |> 
  ggplot(aes(x = outcome, y = count)) +
  geom_col(data = actual_counts, fill = "grey70") +
  stat_pointinterval() +
  labs(x = NULL, y = "Observed and predicted counts",
       title = "Posterior predictions of discrete and continuous outcomes") +
  coord_cartesian(ylim = c(0, 300)) +
  theme_ongo()

actual_middle <- m_full_ordbeta$data |> 
  filter(province_count > 0 & province_count < 1) |> 
  mutate(province_count = prop_to_provinces(province_count))

plot_continuous <- full_posterior |> 
  mutate(.prediction = prop_to_provinces(.prediction)) |> 
  filter(.prediction > 1 & .prediction < 32) |> 
  ggplot(aes(x = .prediction, group = .draw)) +
  geom_density(linewidth = 0.05, bounds = c(1, 32)) +
  geom_density(data = actual_middle, aes(x = province_count),
               inherit.aes = FALSE, bounds = c(1, 32)) +
  labs(x = "Discrete and continuous predicted outcomes",
       y = "Probability density") +
  theme_ongo()
```

```{r combined-posterior-preds, fig.width=6, fig.height=5}
arrow_part <- ggplot() +
  annotate(geom = "segment", x = 0.35, xend = 0.65, y = 0.9, yend = 0.9) +
  annotate(geom = "segment", x = 0.5, xend = 0.5, y = 0.9, yend = 0.1,
           arrow = arrow(angle = 30, type = "closed", length = unit(0.1, "inches"))) +
  scale_x_continuous(limits = c(0, 1)) +
  theme_void()

plot_categories / arrow_part / plot_continuous +
  plot_layout(heights = c(0.45, 0.1, 0.45))
```

```{r}
# ordbetareg's built-in way, just to check
# pp_check_ordbeta(m_full_ordbeta, ndraws = 100)
```

TODO: MCMC diagnostics

TODO: Timing