### 2023-09-13 notebook/model-diagnostics.qmd

---
title: Model diagnostics
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "80%", collapse = TRUE,
  dev = "ragg_png"
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)
```

```{r libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(tidybayes)
library(glue)
library(patchwork)
library(gt)
library(gtExtras)

tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

tar_load(m_full_ordbeta)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ongo())

prop_to_provinces <- function(x, lower = 1, upper = 32) {
  (x * (upper - lower)) + lower
}
```

```{r}
get_variables(m_full_ordbeta)
```

### Traceplots

These should look like hairy caterpillars.

They do.

```{r mcmc-trace}
params_to_show <- c(
  "b_Intercept", "b_work_field_code1Health",
  "phi", "cutzero"
)

m_full_ordbeta |> 
  tidybayes::gather_draws(!!!syms(params_to_show)) |> 
  ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
  geom_line(linewidth = 0.1) +
  scale_color_viridis_d(option = "rocket", end = 0.85) +
  labs(color = "Chain") +
  facet_wrap(vars(.variable), scales = "free_y")
```

### Trace rank plots (trank plots)

These are histograms of the ranks of the parameter draws across the four chains. If the chains are exploring the same space efficiently, the histograms should be similar and overlapping and no one chain should have a specific rank for a long time [@McElreath:2020, 284].

They do.

```{r mcmc-trank}
m_full_ordbeta |> 
  tidybayes::gather_draws(!!!syms(params_to_show)) |> 
  group_by(.variable) |> 
  mutate(draw_rank = rank(.value)) |> 
  ggplot(aes(x = draw_rank, color = factor(.chain))) +
  stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
  scale_color_viridis_d(option = "rocket", end = 0.85) +
  labs(color = "Chain") +
  facet_wrap(vars(.variable), scales = "free_y")
```

## Posterior predictions

The model should generate predictions that align with the observed outcomes. It does.

```{r build-posterior-preds}
full_posterior <- predicted_draws(m_full_ordbeta, 
                                  newdata = m_full_ordbeta$data, 
                                  ndraws = 100)

posterior_categories <- full_posterior |> 
  ungroup() |> 
  mutate(outcome = case_when(
    .prediction == 0 ~ "1",
    .prediction > 0 & .prediction < 1 ~ "1–32",
    .prediction == 1 ~ "32"
  )) |> 
  group_by(outcome, .draw) |> 
  summarize(count = n()) |> 
  ungroup()

actual_counts <- m_full_ordbeta$data |> 
  mutate(outcome = case_when(
    province_count == 0 ~ "1",
    province_count > 0 & province_count < 1 ~ "1–32",
    province_count == 1 ~ "32"
  )) |> 
  group_by(outcome) |> 
  summarize(count = n())

plot_categories <- posterior_categories |> 
  ggplot(aes(x = outcome, y = count)) +
  geom_col(data = actual_counts, fill = "grey70") +
  stat_pointinterval() +
  labs(x = NULL, y = "Observed and predicted counts",
       title = "Posterior predictions of discrete and continuous outcomes") +
  coord_cartesian(ylim = c(0, 300))

actual_middle <- m_full_ordbeta$data |> 
  filter(province_count > 0 & province_count < 1) |> 
  mutate(province_count = prop_to_provinces(province_count))

plot_continuous <- full_posterior |> 
  mutate(.prediction = prop_to_provinces(.prediction)) |> 
  filter(.prediction > 1 & .prediction < 32) |> 
  ggplot(aes(x = .prediction, group = .draw)) +
  geom_density(linewidth = 0.05, bounds = c(1, 32)) +
  geom_density(data = actual_middle, aes(x = province_count),
               inherit.aes = FALSE, bounds = c(1, 32)) +
  labs(x = "Discrete and continuous predicted outcomes",
       y = "Probability density")
```

```{r combined-posterior-preds, fig.width=6, fig.height=5}
arrow_part <- ggplot() +
  annotate(geom = "segment", x = 0.35, xend = 0.65, y = 0.9, yend = 0.9) +
  annotate(geom = "segment", x = 0.5, xend = 0.5, y = 0.9, yend = 0.1,
           arrow = arrow(angle = 30, type = "closed", length = unit(0.1, "inches"))) +
  scale_x_continuous(limits = c(0, 1)) +
  theme_void()

plot_categories / arrow_part / plot_continuous +
  plot_layout(heights = c(0.45, 0.1, 0.45))
```

```{r}
# ordbetareg's built-in way, just to check
# ordbetareg::pp_check_ordbeta(m_full_ordbeta, ndraws = 100)
```