### 2023-09-14 notebook/analysis-issue.qmd
woohoo stacked bar chart with uncertainty in the boundaries
Idea via https://github.com/mjskay/ggdist/issues/109

---
title: "Issue area"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(brms)
library(marginaleffects)
library(tidybayes)
library(ggdist)
library(patchwork)
library(scales)

# Generated via random.org
set.seed(196491)

# Load targets
tar_load(ongo)
tar_load(c(m_full_ordbeta, preds_issue, epreds_issue, issue_indicator_lookup))
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
set_annotation_fonts()

prop_to_provinces <- function(x, lower = 1, upper = 32) {
  (x * (upper - lower)) + lower
}
```

```{r plot-epred-issue}
epreds_issue |> 
  left_join(issue_indicator_lookup, by = join_by(issue_area)) |> 
  mutate(issue_area_nice = fct_reorder(issue_area_nice, .epred)) |> 
  ggplot(aes(x = .epred, y = issue_area_nice)) +
  stat_halfeye(fill = clrs[5]) +
  scale_x_continuous(labels = label_percent(),
                     sec.axis = sec_axis(trans = ~ prop_to_provinces(.),
                                         name = "Expected provinces",
                                         breaks = seq(8, 32, 4))) +
  guides(fill = "none") +
  labs(x = "Expected proportion of provinces", y = NULL) +
  theme_ongo() +
  theme(axis.ticks.x.top = element_line(linewidth = 0.25, color = "grey50"))
```

```{r plot-pred-discrete-continuous-local-bars, fig.width=8, fig.height=3}
preds_issue_plot <- preds_issue |> 
  left_join(issue_indicator_lookup, by = join_by(issue_area)) |> 
  mutate(outcome = case_when(
    .prediction == 0 ~ "1 province",
    .prediction > 0 & .prediction < 1 ~ "1–32 provinces",
    .prediction == 1 ~ "32 provinces"
  )) |> 
  group_by(outcome, issue_area_nice, draw) |> 
  summarize(count = n()) |> 
  group_by(issue_area_nice, draw) |> 
  mutate(prop = count / sum(count))

preds_issue_plot_order <- preds_issue_plot |> 
  filter(outcome == "32 provinces") |>
  group_by(issue_area_nice) |> 
  summarize(median_prop = median(prop)) |> 
  arrange(median_prop) |> 
  mutate(issue_area_nice_labels = str_wrap(issue_area_nice, 15)) |> 
  mutate(issue_area_nice = fct_inorder(issue_area_nice))

preds_issue_text <- preds_issue_plot |> 
  group_by(outcome, issue_area_nice) |> 
  summarize(median_prop = median(prop)) |> 
  group_by(issue_area_nice) |> 
  mutate(y = (median_prop / 2) + lag(cumsum(median_prop), default = 0)) |> 
  mutate(y = 1 - y) |> 
  mutate(prop_nice = label_percent(accuracy = 1)(median_prop)) |> 
  mutate(issue_area_nice = factor(
    issue_area_nice, 
    levels = preds_issue_plot_order$issue_area_nice,
    labels = preds_issue_plot_order$issue_area_nice_labels
  ))

preds_issue_plot |> 
  mutate(issue_area_nice = factor(
    issue_area_nice, 
    levels = preds_issue_plot_order$issue_area_nice,
    labels = preds_issue_plot_order$issue_area_nice_labels
  )) |> 
  ggplot(aes(x = draw, y = prop)) +
  geom_col(aes(fill = outcome), position = "stack", linewidth = 0, width = 1) +
  geom_text(
    data = preds_issue_text,
    aes(x = 250, y = y, label = prop_nice),
    size = 3, fontface = "bold", color = "white"
  ) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  scale_y_continuous(labels = label_percent(), expand = c(0, 0)) +
  scale_fill_manual(values = clrs[c(3, 5, 7)]) +
  facet_wrap(vars(issue_area_nice), strip.position = "bottom", nrow = 1) +
  labs(x = NULL, y = "Proportion of predictions", fill = NULL) +
  theme_ongo() +
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.spacing = unit(5, "pt"),
    strip.text = element_text(vjust = 1, hjust = 0.5, size = rel(0.8), face = "plain")
  )
```