### 2023-12-06 manuscript/appendix.qmd
Make the appendix a separate file
…so that we can have a separate bibliography for software + other methods-related things

---
title: 'Online appendix for "Operational constraints and closing civic space: International NGO restrictions and China’s Overseas NGO law"'
short-title: 'Online appendix for "Operational constraints and closing civic space"'
date: "2023-12-06"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
crossref: 
  tbl-title: Appendix Table
  fig-title: Appendix Figure
---

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(
  echo = FALSE, include = FALSE,
  warning = FALSE, message = FALSE
)

knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "100%", collapse = TRUE
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)
```

# Appendix

## Coding rules for issue ares

1. Education vs. Charity
If charitable activities are all education related, e.g. building schools, donating education facilities and books, coded as education
If the aids provided include other community support, e.g. building bridges, roads, medical facilities, coded as charity

2. Economy and trade vs. Industry association
If the trade promoted is focused on a particular industry (poultry, grains etc.), coded as industry association

3. Arts and culture 
Include sports organizations too.

4. Industry association vs. health, science and technology, arts and culture

There are professional association on health providers and certain techniques. Coded as health, science & technology if work areas go beyond serving members only: for example conducting activities of promoting and training about certain techniques.

If specialized association only serves the member companies: coded as industry associations

5. Health vs. Charity 
If only providing medical aids to underprivileged communities, not building schools etc., coded as health

6. Charity vs. General 
If multiple work, e.g., poverty alleviation, education, health can be encompassed by Charity and humanitarian, then the former

7. Multiple ROs are coded based on the overall work area spans of the INGO, not the specific RO, so to ensure the issue area coding is consistent across ROs

## Modeling details

We use Stan through R and brms to estimate the model. We simulate 4 MCMC chains with 2,000 draws in each chain, 1,000 of which are used for warmup, resulting in 4,000 (1,000 × 4) draws per model parameter. We assess convergence with visual inspection, and all chains converge. Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and model diagnostics—as well as our code and data—are available `ANONYMIZED_URL`.

TODO: Model definition and priors like NVSQ paper